---
title: "Prioritisation Prediction"
author: "UNHCR Data Management & Analysis"
date: "Last updated: `r format(Sys.Date(),  '%d %B %Y')`"
output: 
  html_document:
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_float: yes
---




```{r setup , echo = FALSE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = ""}

#devtools::install_github("cardiomoon/editData")
#devtools::install_github("cardiomoon/ggplotAssist")
#devtools::install_github("calligross/ggthemeassist")
#devtools::install_github('bbc/bbplot')
# install.packages("pacman")

pacman::p_load('dplyr', 'tidyr', 
               'ggplot2',  'ggalt',
               'forcats', 'R.utils', 'png', 
               'grid', 'ggpubr', 'scales',
               'bbplot', 'knitr', 'pander', 'viridis',
               'rgdal', 'cartography','SpatialPosition',
               'OpenStreetMap', 'dummies')

options(scipen = 999)

#Left align text
left_align <- function(plot_name, pieces){
  grob <- ggplot2::ggplotGrob(plot_name)
  n <- length(pieces)
  grob$layout$l[grob$layout$name %in% pieces] <- 2
  return(grob)
}


```





```{r pdataprep , echo = FALSE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = ""}

mainDir <- getwd()
data <- read.csv(paste0(mainDir,"/progrescase-last.csv"), encoding = "UTF-8", na.strings = "")

## Predicted is factor - we check the level - then convert to character to add a factor for NA - and then revert to factor- adding order
#levels(data$predicted.target)
#prop.table(table(data$predicted.target, useNA = "ifany"))
data$predicted.target <- factor(data$predicted.target, levels = c("full.allocation", "reduced.allocation", "no.allocation", "could.not.calculate" ))
#data$predicted.target2 <- as.character(data$predicted.target)
#str(data$predicted.target)
data$predicted.target[is.na(data$predicted.target)] <- "could.not.calculate"

#prop.table(table(data$predicted.target, useNA = "ifany"))

#prop.table(table(data$predicted.target2, useNA = "ifany"))



### Preparing aggaregation per district for mapping purpose

## need to fix some aggregation purpose

## From Zone 1
data$coal5id[data$coal5id == "MAUMBEz001b013"] <- "MAUMBEz001b012"

## From Zone 2
data$coal5id[data$coal5id == "MAUMBEz002b015"] <- "MAUMBEz002b013"

## From Zone 3
data$coal5id[data$coal5id == "MAUMBEz003b011"] <- "MAUMBEz003b010"
data$coal5id[data$coal5id == "MAUMBEz003b012"] <- "MAUMBEz003b010"
data$coal5id[data$coal5id == "MAUMBEz003b013"] <- "MAUMBEz003b010"
data$coal5id[data$coal5id == "MAUMBEz003b014"] <- "MAUMBEz003b010"


##### Subsetting numeric variables
selected.mappoly.numVars <- c( "Num_Inds" ,
                     "Child_0_14",
                     "Youth_15_17",
                     "Work_15_64"  ,
                     "Eldern_65",
                     "Male" ,
                     "Female")
# Create subset of file with observation and selected variables & remove duplicated rows based on IDH
datamappoly1.num <- data[ , c( "coal5id", selected.mappoly.numVars)]

## Convert to numeric variable as they are categoric
#datamappoly.num[,c(selected.mappoly.numVars)] <- lapply(datamappoly.num[,c(selected.mappoly.numVars)], numeric)

## aggregate numeric value  based on mean
datamappoly1.num2 <- datamappoly1.num %>%
  group_by(coal5id) %>%
  summarise_all(funs(mean(., na.rm = TRUE))) %>%
  mutate_if(is.numeric, funs(round(., 2)))

datamappoly1.num2 <- as.data.frame(datamappoly1.num2)


## Getting a count
datamappoly1.n <-  data[ ,c("coal5id", "Num_Inds")] %>% group_by(coal5id) %>% mutate(count = n())
datamappoly1.n <- as.data.frame(unique(datamappoly1.n[ ,c('coal5id', 'count')]))


##### Subsetting categoric variables
selected.mappoly.catVars <- "predicted.target"
# Create subset of file with observation and selected variables & remove duplicated rows based on IDH
datamappoly1.cat <- data[ , c( 'coal5id', selected.mappoly.catVars)]

## Convert to factor variable as they are categoric
#datamappoly1.cat[,c(selected.mappoly.catVars)] <- lapply(datamappoly1.cat[,c(selected.mappoly.catVars)], factor)
## Hot coding variable
datamappoly1.cat <- dummy.data.frame(datamappoly1.cat, names = selected.mappoly.catVars, sep = ".")

## aggregate numeric value  based on mean
datamappoly1.cat2 <- datamappoly1.cat %>%
  group_by(coal5id) %>%
  summarise_all(funs(sum)) #%>%
  #mutate_if(is.numeric, funs(round(., 2)))

### Renaming select multiple variable
#for (i in 1:ncol(datamappoly1.cat2)) {
#  if (names(datamappoly1.cat2)[i] %in% as.character(selected.mappoly.cat.multi$fullname2)) {
#    names(datamappoly1.cat2)[i] = as.character(selected.mappoly.cat.multi$fullname[match(names(datamappoly1.cat2)[i], selected.mappoly.cat.multi$fullname2)])
#  }
#}

datamappoly1.cat2 <- as.data.frame(datamappoly1.cat2)

## Bind everything
datamappoly1 <- merge(x = datamappoly1.n, y = datamappoly1.cat2, by = "coal5id")
for (i in 3:ncol(datamappoly1)) { datamappoly1[i] <- round(datamappoly1[i]/datamappoly1$count,2) }

datamappoly1 <- merge(x = datamappoly1, y = datamappoly1.num2, by = "coal5id")

rm(datamappoly1.n,  datamappoly1.cat, datamappoly1.cat2,  datamappoly1.num, datamappoly1.num2)

## getting correct district and gov from coordinates
campdistrict <- readOGR(paste0(mainDir,"/campdistrict.geojson"), verbose = FALSE)


## Fortify
campdistrict.fort <- fortify(campdistrict, region = "coal5id")
datamappoly1 <- datamappoly1[!(is.na(datamappoly1$coal5id)), ]
datamappoly1$id <- datamappoly1$coal5id
campdistrict.map.fort <- merge( x = campdistrict.fort, y = datamappoly1, by = "id")

rm(campdistrict.fort)

idList <- unique(campdistrict.map.fort$id)
centroids.df <- as.data.frame(coordinates(campdistrict))
names(centroids.df) <- c("Longitude", "Latitude")

centroids.df <- data.frame(id = idList, centroids.df)
centroids.df <- merge( x = centroids.df, y = datamappoly1, by = "id")


## get extend
xmin <- as.data.frame(campdistrict@bbox)[1,1]
xmax <- as.data.frame(campdistrict@bbox)[1,2]
ymin <- as.data.frame(campdistrict@bbox)[2,1]
ymax <- as.data.frame(campdistrict@bbox)[2,2]


## Extend the extend
xmin <- xmin - ((xmax - xmin)/10)
xmax <- xmax + ((xmax - xmin)/10)
ymin <- ymin - ((ymax - ymin)/10)
ymax <- ymax + ((ymax - ymin)/10)


## Map background
map <- openmap(c(lat = ymax, lon = xmin ), c(lat = ymin, lon = xmax), type = "osm")
mapLatLon <- openproj(map)

#names(campdistrict.map.fort)



```


This reports is automatically generated and presents the results of the assistance prioritisation predictive model.

The analysis was performed on `r nrow(data)` case-level records. Among all those cases, a certain number does not have sufficient information (missing variables) to perfrom the calculation.


The prediction are presented using the best predictive model for categorical prediction ([Random Decision Forest](https://en.wikipedia.org/wiki/Random_forest) )



# Results of predictive model

## By categories


```{r plot1, echo = FALSE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = ""}

plot1 <- ggplot(data, aes(x = predicted.target)) +
  geom_bar( stat = "count",  
           fill = "#1380A1") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  geom_label(aes(label = ..count.., y = ..count..), stat = "count",
             hjust = 1, 
             vjust = 0.5, 
             colour = "white", 
             fill = NA, 
             label.size = NA, 
             family = "Helvetica", 
             size = 6) +
  coord_flip() +
  labs(title = "How many per category?",  subtitle = "Predicted category by cases #", caption = "UNHCR Mauritania", x = NULL, y = NULL) +
  bbc_style() +
  theme( plot.title = element_text(size = 13),
         plot.subtitle = element_text(size = 11),
         plot.caption = element_text(size = 7, hjust = 1),
         axis.text = element_text(size = 10),
         strip.text.x = element_text(size = 11))

ggpubr::ggarrange(left_align(plot1, c("subtitle", "title")), ncol = 1, nrow = 1)




```

## By probabilities

```{r plot2, echo = FALSE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = ""}

plot2 <- ggplot(data, aes(x = predicted.prob)) +
  geom_bar( stat = "count", fill = "#1380A1",  width = .8) +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  
 facet_wrap(~  predicted.target, ncol = 2) +
 # coord_flip() +
  scale_y_continuous(limits = c(0, 50)) +
 # scale_x_continuous( breaks=NULL) +
  labs(title = "According to Predicted probability",  subtitle = "Predicted category by cases #", caption = "UNHCR Mauritania", x = NULL, y = NULL) +
  bbc_style() +
  theme( plot.title = element_text(size = 13),
         plot.subtitle = element_text(size = 11),
         plot.caption = element_text(size = 7, hjust = 1),
         axis.text = element_text(size = 10),
         strip.text.x = element_text(size = 11))

ggpubr::ggarrange(left_align(plot2, c("subtitle", "title")), ncol = 1, nrow = 1)



```


## By Case size

```{r plot3, echo = FALSE,   warning = FALSE, tidy = TRUE, message=FALSE, comment = ""}

plot3 <- ggplot(data, aes(x = Case.size)) +
  geom_bar( stat = "count",  
           fill = "#1380A1") +
  geom_hline(yintercept = 0, size = 1, colour = "#333333") +
  
  facet_wrap(~  predicted.target, ncol = 2) +
  geom_label(aes(label = ..count.., y = ..count..), stat = "count",
             hjust = 1, 
             vjust = 0.5, 
             colour = "white", 
             fill = NA, 
             label.size = NA, 
             family = "Helvetica", 
             size = 6) +
  coord_flip() +
  labs(title = "According to Case size?",  subtitle = "Predicted category by cases #", caption = "UNHCR Mauritania", x = NULL, y = NULL) +
  bbc_style() +
  theme( plot.title = element_text(size = 13),
         plot.subtitle = element_text(size = 11),
         plot.caption = element_text(size = 7, hjust = 1),
         axis.text = element_text(size = 10),
         strip.text.x = element_text(size = 11))

ggpubr::ggarrange(left_align(plot3, c("subtitle", "title")), ncol = 1, nrow = 1)



```



# Camp maps

## Overview of the total registered population

```{r count, echo=FALSE, warning=FALSE, cache=FALSE, tidy = TRUE, message=FALSE, comment = "", fig.height= 4,size="small"}

autoplot(mapLatLon) +
  geom_polygon(data = campdistrict.map.fort,
               aes(x = long, y = lat,  group = group),
               colour = "white", alpha = 0.6 ) +
  geom_point(data = centroids.df,
               aes(x = Longitude, y = Latitude, 
                   size = count )) +
  geom_label(data = centroids.df,
               aes(x = Longitude, y = Latitude, label = count), 
            # hjust = 1, 
           #  vjust = 0.5, 
             colour = "white", 
             fill = NA, 
             label.size = NA, 
             family = "Helvetica", 
             size = 3) +
  
  coord_equal() +
  scale_size_continuous( breaks = c(0, 50, 200, 300, 400, 550, 1000),
                         trans =  "log10", #label =c(less than 50)
                      name = "Number of cases for that modality",
                      guide = guide_legend( )) +
  labs(title = "How is population spread in the camp?",
       subtitle = "Cases # by block") +
  
  labs( title = "How is population spread in the camp?" , subtitle = "# of Cases by block", caption = "UNHCR Mauritania", x = NULL, y = NULL) +
  
  bbc_style() +
  theme(legend.position = "none",
        axis.line = element_blank(),
         axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         axis.title.x = element_blank(),
         axis.title.y = element_blank(),
         legend.text.align = 0, 
         legend.background = element_rect(fill = alpha('white', 0.0)),
         legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"),  
         legend.title = element_text(size = 6), 
         plot.title = element_text(size = 13),
         plot.subtitle = element_text(size = 11),
         plot.caption = element_text(size = 7)) 



```


## Prediction per categories

```{r no.allocation, echo=FALSE, warning=FALSE, cache=FALSE, tidy = TRUE, message=FALSE, comment = "", fig.height= 4, size="small"}

autoplot(mapLatLon) +
  geom_polygon(data = campdistrict.map.fort,
               aes(x = long, y = lat, 
                   fill = round(predicted.target.no.allocation,2) , group = group),
               colour = "white", alpha = 0.5 ) +
  geom_point(data = centroids.df,
               aes(x = Longitude, y = Latitude, 
                   size = predicted.target.no.allocation * count ), colour = "white") +
  
  geom_label(data = centroids.df,
               aes(x = Longitude, y = Latitude, label = round(predicted.target.no.allocation * count,0)), 
             colour = "black", 
             fill = NA, 
             label.size = NA, 
             family = "Helvetica", 
             size = 4) +
  
  coord_equal() +
  scale_fill_viridis(
                      name = "Percentage for that modality", labels = percent,
                      guide = guide_legend( direction = "vertical", label.position = "right",
                                            keyheight = unit(2, units = "mm"),  
                                            keywidth = unit(length(labels)*10, units = "mm"),
                                            title.position = 'top',  title.hjust = 0.5, label.hjust = 1#, 
                                            #nrow = 1, byrow = T, reverse = T )
                                            )) +

    scale_size_area( name = "Number of cases for that modality",  trans =  "log10", guide = FALSE  ) +
  
  labs( title = "No allocation" , subtitle = "Predicted Prioritisation", caption = "UNHCR Mauritania", x = NULL, y = NULL) +
  
  bbc_style() +
  theme( legend.position = "right", 
         legend.text.align = 0, 
         legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"), 
         legend.title = element_text(size = 8), 
         panel.border = element_blank(), 
         plot.title = element_text(size = 13),
         plot.subtitle = element_text(size = 11),
         plot.caption = element_text(size = 7, hjust = 1),
         axis.line = element_blank(),
         axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         axis.title.x = element_blank(),
         axis.title.y = element_blank())  


```



```{r reduced.allocation2, echo=FALSE, warning=FALSE, cache=FALSE, tidy = TRUE, message=FALSE, comment = "", fig.height= 4,size="small"}

autoplot(mapLatLon) +
  geom_polygon(data = campdistrict.map.fort,
               aes(x = long, y = lat, 
                   fill = round(predicted.target.reduced.allocation,2) , group = group),
               colour = "white", alpha = 0.5 ) +
  geom_point(data = centroids.df,
               aes(x = Longitude, y = Latitude, 
                   size = predicted.target.reduced.allocation * count ), colour = "white") +
  
  geom_label(data = centroids.df,
               aes(x = Longitude, y = Latitude, label = round(predicted.target.reduced.allocation * count,0)), 
             colour = "black", 
             fill = NA, 
             label.size = NA, 
             family = "Helvetica", 
             size = 4) +
  
  coord_equal() +
  scale_fill_viridis(
                      name = "Percentage for that modality", labels = percent,
                      guide = guide_legend( direction = "vertical", label.position = "right",
                                            keyheight = unit(2, units = "mm"),  
                                            keywidth = unit(length(labels)*10, units = "mm"),
                                            title.position = 'top',  title.hjust = 0.5, label.hjust = 1#, 
                                            #nrow = 1, byrow = T, reverse = T )
                                            )) +

    scale_size_area( name = "Number of cases for that modality",trans =  "log10", guide = FALSE  ) +
  
  labs( title = "Reduced allocation" , subtitle = "Predicted Prioritisation", caption = "UNHCR Mauritania", x = NULL, y = NULL) +
  
  bbc_style() +
  theme( legend.position = "right", 
         legend.text.align = 0, 
         legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"), 
         legend.title = element_text(size = 8), 
         panel.border = element_blank(), 
         plot.title = element_text(size = 13),
         plot.subtitle = element_text(size = 11),
         plot.caption = element_text(size = 7, hjust = 1),
         axis.line = element_blank(),
         axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         axis.title.x = element_blank(),
         axis.title.y = element_blank()) 



```



```{r full.allocation, echo=FALSE, warning=FALSE, cache=FALSE, tidy = TRUE, message=FALSE, comment = "", fig.height= 4,size="small"}

autoplot(mapLatLon) +
  geom_polygon(data = campdistrict.map.fort,
               aes(x = long, y = lat, 
                   fill = round(predicted.target.full.allocation,2) , group = group),
               colour = "white", alpha = 0.5 ) +
  geom_point(data = centroids.df,
               aes(x = Longitude, y = Latitude, 
                   size = predicted.target.full.allocation * count ), colour = "white") +
  
  geom_label(data = centroids.df,
               aes(x = Longitude, y = Latitude, label = round(predicted.target.full.allocation * count,0)), 
             colour = "black", 
             fill = NA, 
             label.size = NA, 
             family = "Helvetica", 
             size = 4) +
  
  coord_equal() +
  scale_fill_viridis(
                      name = "Percentage for that modality", labels = percent,
                      guide = guide_legend( direction = "vertical", label.position = "right",
                                            keyheight = unit(2, units = "mm"),  
                                            keywidth = unit(length(labels)*10, units = "mm"),
                                            title.position = 'top',  title.hjust = 0.5, label.hjust = 1#, 
                                            #nrow = 1, byrow = T, reverse = T )
                                            )) +

    scale_size_area( name = "Number of cases for that modality",trans =  "log10", guide = FALSE  ) +
  
  labs( title = "Full allocation" , subtitle = "Predicted Prioritisation", caption = "UNHCR Mauritania", x = NULL, y = NULL) +
  
  bbc_style() +
  theme( legend.position = "right", 
         legend.text.align = 0, 
         legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"), 
         legend.title = element_text(size = 8), 
         panel.border = element_blank(), 
         plot.title = element_text(size = 13),
         plot.subtitle = element_text(size = 11),
         plot.caption = element_text(size = 7, hjust = 1),
         axis.line = element_blank(),
         axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         axis.title.x = element_blank(),
         axis.title.y = element_blank()) 


```




```{r could.not.calculate, echo=FALSE, warning=FALSE, cache=FALSE, tidy = TRUE, message=FALSE, comment = "", fig.height= 4,size="small"}

autoplot(mapLatLon) +
  geom_polygon(data = campdistrict.map.fort,
               aes(x = long, y = lat, 
                   fill = round(predicted.target.could.not.calculate,2) , group = group),
               colour = "white", alpha = 0.5 ) +
  geom_point(data = centroids.df,
               aes(x = Longitude, y = Latitude, 
                   size = predicted.target.could.not.calculate * count ), colour = "white") +
  
  geom_label(data = centroids.df,
               aes(x = Longitude, y = Latitude, label = round(predicted.target.could.not.calculate * count,0)), 
             colour = "black", 
             fill = NA, 
             label.size = NA, 
             family = "Helvetica", 
             size = 4) +
  
  coord_equal() +
  scale_fill_viridis(
                      name = "Percentage for that modality", labels = percent,
                      guide = guide_legend( direction = "vertical", label.position = "right",
                                            keyheight = unit(2, units = "mm"),  
                                            keywidth = unit(length(labels)*10, units = "mm"),
                                            title.position = 'top',  title.hjust = 0.5, label.hjust = 1#, 
                                            #nrow = 1, byrow = T, reverse = T )
                                            )) +

    scale_size_area( name = "Number of cases for that modality",trans =  "log10", guide = FALSE  ) +
  
  labs( title = "Could not calculate" , subtitle = "Predicted Prioritisation", caption = "UNHCR Mauritania", x = NULL, y = NULL) +
  
  bbc_style() +
  theme( legend.position = "right", 
         legend.text.align = 0, 
         legend.text = element_text(size = 7, hjust = 0, color = "#4e4d47"), 
         legend.title = element_text(size = 8), 
         panel.border = element_blank(), 
         plot.title = element_text(size = 13),
         plot.subtitle = element_text(size = 11),
         plot.caption = element_text(size = 7, hjust = 1),
         axis.line = element_blank(),
         axis.text.x = element_blank(),
         axis.text.y = element_blank(),
         axis.ticks = element_blank(),
         axis.title.x = element_blank(),
         axis.title.y = element_blank()) 
```











